\cleardoublepage
%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

% \begin{flushright}{\slshape    
%     We have seen that computer programming is an art, \\ 
%     because it applies accumulated knowledge to the world, \\ 
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- {Donald E. Knuth}%\defcitealias{knuth:1974}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
The work presented in this thesis would not have been possible without the help of my supervisors, co-workers, colleagues, and friends.

First, I would like to thank my first supervisor Prof. Loe Feijs, my second supervisor Prof. Panos Markopoulos, and my co-supervisor Dr. Jun Hu. I would also like to thank the members of my reading committee: Prof. Harold Thimbleby, Dr. Stefan Wensveen and Prof. Johan Lukkien. Thank you for all your help, suggestions and guidance.

Bram van der Vlist --- who was my teammate on this adventure --- without you this would definitely not have been possible.

I would like to thank Aly Syed, Riccardo Trevisan, Sriram Srinivasan, Hans van Amstel, Stefan Rapp, Sachin Bhardwaj and Tanir Ozcelebi for their valuable contributions to the smart home pilot.

While working on the software tools and ontologies described in this thesis, Holger Knublauch and Scott Henninger from TopQuadrant were always ready to provide me with assistance.

I will always remember my PhD colleagues who I shared an office with at the university: Alex Juarez, Sibrecht Bouwstra, Roman Gorbunov, Jan Gillesen, Misha Croes and Marija Nakevska. You guys made my three years at TU Eindhoven very enjoyable and a lot of fun. Also a sincere thank you to Ellen Konijnenberg for helping me with all the organisational issues. 

I would like to thank my parents and my brother: \emph{Ma, Pa en Riaan - baie dankie vir julle ondersteuning and dat julle in my glo.} This thesis is dedicated to my fianc\'ee --- Margriet van de Goor, \emph{ik hou van je}!

\endgroup



