\cleardoublepage
\pagestyle{empty}

%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{publications}
\chapter*{Publications related to this research}

Some ideas and figures have appeared previously in the following publications:

\bigskip

1. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Rapp, S., Hu, J., \& Feijs, L.M.G.  (under review). Configuring and controlling ubiquitous computing infrastructure with semantic connections: a tangible and an AR approach. \emph{Personal and Ubiquitous Computing.} 29 pages. \\

\noindent 2. \textbf{Niezen, G.}, Vlist, B.J.J. van der, Hu, J., \& Feijs, L.M.G.  (under review). Semantic Connections Theory: Enabling Interaction Designers and Developers to Create Interoperable Smart Objects. \emph{ACM Transactions on Interactive Intelligent Systems (TiiS).} 32 pages. \\

\noindent 3. Peeters, J. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L.M.G. (2012). A Study on a Tangible Interaction Approach to Managing Wireless Connections in a Smart Home Environment.  In L.- L. Chen, T. Djajadiningrat, L. Feijs, S. Fraser, S. Kyffin, \&  D. Steffen (Eds.) \emph{7th International Workshop on the Design \& Semantics of Form \& Movement (DeSForM) 2012.} pages 187--196. Wellington, New Zealand: Koninklijke Philips Electronics N.V.\\

\noindent 4. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Rapp, S., Hu, J., \& Feijs, L.M.G. (2012). Controlling Smart Home Environments with Semantic Connections: a Tangible and an AR Approach. In L.- L. Chen, T. Djajadiningrat, L. Feijs, S. Fraser, S. Kyffin, \&  D. Steffen (Eds.) \emph{7th International Workshop on the Design\& Semantics of Form \& Movement (DeSForM) 2012.} pages 160--169.  Wellington, New Zealand: Koninklijke Philips Electronics N.V.\\

\noindent 5. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L.M.G. (2012). Semantic Connections: a New Interaction Paradigm for Smart Environments. In L.- L. Chen, T. Djajadiningrat, L. Feijs, S. Fraser, S. Kyffin, \&  D. Steffen (Eds.) \emph{7th International Workshop on the Design \& Semantics of Form \& Movement (DeSForM) 2012.} pages 16--26. Wellington, New Zealand: Koninklijke Philips Electronics N.V.\\

\noindent 6. \textbf{Niezen, G.}, Vlist, B.J.J. van der, Bhardwaj, S. \& Ozcelebi, T. (2012). Performance Evaluation of a Semantic Smart Space Deployment. \emph{4rd International Workshop on Sensor Networks and Ambient Intelligence (SeNAmI 2012).} March 19--23, 2012, Lugano, Switzerland.\\

\noindent 7. Hu, J., Vlist, B.J.J. van der, \textbf{Niezen, G.}, Willemsen, W., Willems, D. \& Feijs, L. Designing the Internet of Things for Learning Environmentally Responsible Behaviour. \emph{Interactive Learning Environments}, Taylor \& Francis.\\

\noindent 8. \textbf{Niezen, G.}, Vlist, B.J.J. van der, Hu, J. \& Feijs, L.M.G. (2012). Using semantic transformers to enable interoperability between media devices in a ubiquitous computing environment. In Rautiainen, M. et al. (Eds.), \emph{Grid and Pervasive Computing Workshops, Lecture Notes in Computer Science, Vol. 7096.} pages 44--53. Springer Berlin / Heidelberg.\\

\noindent 9. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L. (2011). Design semantics of connections in a smart home environment. \emph{Creation and Design, 13}(2) p.p. 18--24.\\

\noindent 10. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L.M.G. (2011) \emph{Interaction Primitives: Describing Interaction Capabilities of Smart Objects in Ubiquitous Computing Environments.} IEEE Africon 2011, September 13--15, Livingstone, Zambia.\\

\noindent 11. Kwak, M., \textbf{Niezen, G.}, Vlist, B. van der, Hu, J., \& Feijs, L. (2011). \emph{Tangible interfaces to digital connections, centralized versus de- centralized.} In Z. Pan, A. Cheok, W. Mueller, and X. Yang (Eds.), Transactions on edutainment V (Vol. 6530, p. 132--146). Springer Berlin / Heidelberg.\\

\noindent 12. Hu, J., Vlist, B.J.J. van der, \textbf{Niezen, G.}, Willemsen, W., Willems, D. \& Feijs, L. (2011) \emph{ Learning from Internet of Things for Improving Environmentally Responsible Behavior.} In M. Chang et al. (Eds.), Edutainment 2011, LNCS Vol. 6872, pp. 292--299, Springer-Verlag, Berlin / Heidelberg.\\

\noindent 13. Willemsen, W., Hu, J., \textbf{Niezen, G.} \& Vlist, B.J.J. van der (2011). \emph{Using Game Elements to Motivate Environmentally Responsible Behaviour.} Proceedings of IADIS Game and Entertainment Technologies 2011 (GET 2011), Rome.\\

\noindent 14. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L.M.G. (2010). \emph{Design semantics of connections in a smart home environment.} In L.- L. Chen, T. Djajadiningrat, L. Feijs, S. Kyffin, D. Steffen, \&  B. Young (Eds.), \emph{Proceedings of design and semantics of form and movement (DeSForM) 2010} (p. 48--56). Lucerne, Switzer-land: Koninklijke Philips Electronics N.V.\\

\noindent 15. Vlist, B.J.J. van der, \textbf{Niezen, G.}, Hu, J., \& Feijs, L.M.G. (2010). \emph{Semantic connections: Exploring and manipulating connections in smart spaces.} In Computers and Communications (ISCC), 2010 IEEE Symposium on (pp. 1--4). Riccione, Italy: IEEE.\\

\noindent 16. \textbf{Niezen, G.}, Vlist, B.J.J. van der, Hu, J. \& Feijs, L.M.G. (2010). \emph{From events to goals: supporting semantic interaction in smart environments.} Proceedings of the Computers and Communications (ISCC), 2010 IEEE Symposium on, (pp. 1029--1034). IEEE.\\


% \bibliographystylemypub{plainnat}
% \bibliographymypub{MyPublications}
% \nocite{*}